Planning with Hierarchical Temporal Memory for deterministic Markov Decision Problem

### 2020-10-29 reports/v1_0/paper_eng/paper_eng.tex
fix citations

\begin{abstract}
  Among the key problems in artificial intelligence is sequential decision making, which can be formalized as Markov Decision Process (MDP). One approach to solve it, called model-based Reinforcement Learning (RL), combines learning the model of the environment and the global policy. Having a good model of the environment opens up such properties as data efficiency and targeted exploration.

@@

Model-based RL can address issues of both planning and model-free RL by first learning the environment dynamics, and then to exploit it for more efficient learning of the agent's policy (\cite{sutton_barto_2018}). Typically, the model is represented by MDP and consists of two components: a state transition model and a reward model. After the model has been learned, MDP planning method such as MCTS (\cite{Coulom_2007_mcts}) can be applied to infer the optimal policy. Many model-based RL approaches build a model that operates at the pixel-level (\cite{kaiser_2020_modelbased}), which is hungry for computational resources and attentive to unimportant details. Another approach is to try building a latent-space model. For example, a latent-space model can represent an abstract MDP that's equivalent to the real one in terms of state values, i.e. the cumulative reward of the planned trajectory (\cite{silver_2020_muzero}).

To model the environment it's common to use ANNs (\cite{silver_2020_muzero,Ha_Schmidhuber_2018_worldmodels}). It's also very tempting to take more inspiration from human brains, especially for how our memory functions (\cite{Hassabis_2017_neuro}). One example of such human-like memory model is Hierarchical Temporal Memory (HTM) framework (\cite{George_Hawkins_2009}). This framework utilizes discrete binary sparse distributed representations (SDR) (\cite{ahmad_2016_sdr}), which enables more efficient data processing, high noise tolerance and a natural way to set similarity metric between data embeddings. It also provides a model of the sequential temporal memory (TM) (\cite{hawkins_TM}). Learning in TM isn't based on the gradient descend, and typically it learns with a faster rate compared to common ANNs.

@@

>>>>

\begin{abstract}
  Sequential decision making is among the key problems in Artificial Intelligence. It can be formalized as Markov Decision Process (MDP). One approach to solve it, called model-based Reinforcement Learning (RL), combines learning the model of the environment and the global policy. Having a good model of the environment opens up such properties as data efficiency and targeted exploration.

@@

Model-based RL can address issues of both planning and model-free RL by first learning the environment dynamics, and then to exploit it for more efficient learning of the agent's policy (\cite{moerland_2020_modelbased}). Typically, the model is represented by MDP and consists of two components: a state transition model and a reward model. After the model has been learned, MDP planning method such as MCTS (\cite{Coulom_2007_mcts}) can be applied to infer the optimal policy. Many model-based RL approaches build a model that operates at the pixel-level (\cite{kaiser_2020_modelbased}), which is hungry for computational resources and attentive to unimportant details. Another approach is to try building a latent-space model. For example, a latent-space model can represent an abstract MDP that's equivalent to the real one in terms of state values, i.e. the cumulative reward of the planned trajectory (\cite{silver_2020_muzero}).

To model the environment it's common to use ANNs (\cite{silver_2020_muzero,Ha_Schmidhuber_2018_worldmodels}). It's also very tempting to take more inspiration from human brains, especially for how our memory functions (\cite{Hassabis_2017_neuro}). One example of such human-like memory model is Hierarchical Temporal Memory (HTM) framework (\cite{George_Hawkins_2009}). This framework utilizes discrete binary sparse distributed representations (SDR) (\cite{Cui_Ahmad_Hawkins_2017_sdr}), which enables more efficient data processing, high noise tolerance and a natural way to set similarity metric between data embeddings. It also provides a model of the sequential temporal memory (TM) (\cite{hawkins_TM}). Learning in TM isn't based on the gradient descend, and typically it learns with a faster rate compared to common ANNs.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{agent_scheme}
  \caption{Schematic view of our agent and its training loop.} \label{fig_agent_scheme}
\end{figure}
